To generate the Mandelbrot set, we use the \textbf{Escape Time Algorithm}.

\begin{algorithm}[H]
    \SetAlgoLined
    \caption{Escape Time Algorithm}
    \label{alg:escape_time_algorithm}
    \KwData{Complex number $c$, maximum number of iterations $N$}
    \KwResult{Number of iterations $n$}
    z = 0\;
    \While{$|z| \leq 2$ and $n < N$}{
        $z \leftarrow z^2 + c$\;
        $n \leftarrow n + 1$\;
    }
    \Return $n$\;
\end{algorithm}

The iteration threshold $N$ is conventionally configured at $N = 1000$
, which establishes a balance between computational efficiency and visual fidelity.
Our implementation employs this parameterization scheme for fractal generation.

To enhance the structural analysis, we introduce a boundary detection methodology based on the gradient field of the
layer function.
This approach enables the reconstruction of high-fidelity visualizations preserving the set's topological features.
Specifically, we use the Sobel operator\textsuperscript{\cite{sobel1968isotropic}}
to compute the discrete gradient approximation.

A Sobel operator is defined as follows:

\begin{equation}
    \label{eq:sobel_operator}
    G_x = \begin{bmatrix}
              -1 & 0 & 1 \\
              -2 & 0 & 2 \\
              -1 & 0 & 1
    \end{bmatrix}
    \quad
    G_y = \begin{bmatrix}
              -1 & -2 & -1 \\
              0  & 0  & 0  \\
              1  & 2  & 1
    \end{bmatrix}
\end{equation}

Then, the gradient magnitude is computed as follows:

\begin{equation}
    \label{eq:gradient_magnitude}
    \parallel \nabla L \parallel \ = \sqrt{(G_x * L)^2 + (G_y * L)^2}
\end{equation}

However, in practice, we approximate the gradient magnitude using the absolute value of the gradient components, and
we take the arithmetic mean of the two components.
This simplification allows for a more computationally efficient.

Then, we set \textit{GradientThreshold}
to 0.5, and pick out the pixels with gradient magnitude higher than the threshold.
These pixels are likely to be on the edge of the Mandelbrot set, so we can automatically detect the boundary of the
set.

Here is the implementation:

\begin{lstlisting}[label={lst:detect_high_gradient}, gobble=8]
    cv::Mat detectHighGradient(const cv::Mat &matrix) {
        constexpr static double GRADIENT_THRESHOLD = 0.5;

        cv::Mat normalized;
        cv::normalize(matrix, normalized, 0, 255, cv::NORM_MINMAX, CV_8UC1);

        cv::Mat grad_x, grad_y;
        cv::Sobel(normalized, grad_x, CV_32F, 1, 0);
        cv::Sobel(normalized, grad_y, CV_32F, 0, 1);

        cv::Mat abs_grad_x, abs_grad_y, grad_mag;
        cv::convertScaleAbs(grad_x, abs_grad_x);
        cv::convertScaleAbs(grad_y, abs_grad_y);
        cv::addWeighted(abs_grad_x, 0.5, abs_grad_y, 0.5, 0, grad_mag);

        double min_val, max_val;
        cv::minMaxLoc(grad_mag, &min_val, &max_val);
        double threshold = min_val + (max_val - min_val) * GRADIENT_THRESHOLD;

        cv::Mat mask = grad_mag;
        cv::threshold(grad_mag, mask, threshold, 255, cv::THRESH_BINARY);
        return mask;
    }
\end{lstlisting}

With the mask, we can detect the area with high gradient magnitude, which is likely to be the fractal boundary.
In detail, we split the matrix with several blocks, and for each block, we calculate the total number of high
gradient pixels.
Then, we can determine the block with the most high gradient pixels, and we can zoom in on this area to get a more
detailed image of the boundary.
Here is the pseudocode:


\begin{algorithm}[H]
    \SetAlgoLined
    \caption{Detect Boundary}
    \label{alg:detect_boundary}
    \KwData{Escape time matrix $matrix$, block rows and columns count $n$}
    \KwResult{Boundary area}
    $mask \leftarrow detectHighGradient(matrix)$\;
    $maxHighGradientPixels \leftarrow 0$\;
    $maxBlock \leftarrow null$\;
    Split matrix into $n \times n$ blocks\;
    \For{each block}{
        Count the number of high gradient pixels\;
        \If{high gradient pixels > maxHighGradientPixels}{
            $maxHighGradientPixels \leftarrow high gradient pixels$\;
            $maxBlock \leftarrow block$\;
        }
    }
    \Return maxBlock\;
\end{algorithm}

We pick out the maximum block and set its center as the new center of the image.
Then, we zoom in on this area to get a more detailed image of the boundary.
To implement this, we need to adjust the scale of the image and the center of the image.
OpenCV provides several functions to generate the scale matrix, and we can use the \textit{warpAffine}
function to adjust the image.
These code are in \href{https://github.com/AI1379/MandelbrotSet/blob/master/src/VideoGenerator.h}
{\texttt{src/VideoGenerator.h}}.